---
title: "Homework 2"
author: "Michael Ahn"
date: "September 23, 2017"
output:
  pdf_document: default
  html_document: default
---

  ```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Problem 1: Insects

### a.) By Hand
  Characteristic A supposedly has a 0.2 probability of appearing in insects. We have 18 insects, 0 of which contain characteristic A.
  H0: p = 0.2  
  H1: p ~= 0.2  
  We thus have the following...
```{r}
p <- 0.2
n <- 18
t <- 0
```
 
  Since t < n*p we use the rhs for a two-tailed test
```{r}
rhs <- seq(ceiling(n*p), n)
tU <- sum(dbinom(rhs, n, p) <= dbinom(t, n, p))
  # use tU value to find p-value
phi <- pbinom(t, n, p) + pbinom(n - tU, n, p, lower.tail = FALSE)
phi.approx <- 2*pbinom(t, n, p)
c(phi, phi.approx)
```
  with a p-value of ~0.034, we reject the null hypothesis and conclude that the probability is different

### b.) Using Software Library
```{r insect}
binom.test(t, n, p)
```  
  with a p-value of 0.03429, we come to the same conclusion using the software library function

## Problem 2: Dice

### a.) By Hand
Dice is thrown 180 times, 'seven' is rolled 38 times. Probability of 'seven' is is p = 1/6 for fair dice.  
H0: p = 1/6  
H1: p > 1/6

Since 38 is greater than 180/6, we assert the alternative hypothesis that the probability is greater than 1/6...
```{r}
n <- 180
t <- 38
p <- 1/6

1 - pbinom(t-1, n,p)
```
Assuming a significance level of 0.05, we cannot reject the null hypothesis, and must conclude that the dice is still fair

### b.) Using Software Library
```{r}
binom.test(t, n, p, alternative = "greater")
```
Again, if we go by a significance level of 0.05, given a p value of .6997, we cannot reject the null hypothesis. This p-value stands for our confidence or probability that the data supports the null hypothesis, and it is not below the rejection level.

## Problem 3: Weight

### 1. Testing the Hypothesis (weight = 103)
Given the data, we check whether the median weight is really 103  
H0: weight = 103  
H1: weight ~= 103  

#### a.) By Hand
If we try to test this hypothesis by finding a 90% confidence interval...
```{r}
weight <- c(142, 134, 98, 119, 131, 103, 154, 122, 93, 137, 86, 119, 161, 144, 158, 165, 81, 117, 128, 103)

alpha <- 0.05
n <- length(weight)
p <- 0.5

r <- qbinom(alpha/2, n, p)
s <- qbinom(1 - alpha/2, n, p)
weight.sorted <- sort(weight)

c(weight.sorted[r], weight.sorted[s])
```
This gives us the approximated confidence interval.
```{r}
alpha1 <- pbinom(r-1, n, p)
alpha2 <- 1 - pbinom(s-1, n, p)
1 - alpha1 - alpha2
```
We are actually only 92% confident that the median falls in the interval [103, 137] so we do not reject the null hypothesis.
Anther approximation...

```{r}
t <- 103
p <- 0.5
n <- length(weight)

t1 <- n - sum(weight > t)
t2 <- n - sum(weight >= t)

phi <- 2*min(pbinom(t1, n, p), 1 - pbinom(t2-1, n, p))
phi
```
Cannot reject null hypothesis using a 95% significance.

#### b.) Software Library  
```{r}
binom.test(t1,n,p)
binom.test(t2,n,p)
```
The binom.test gives us a pvalue of ~0.11 which is still not enough to reject the null hypothesis given a 95% confidence, or even a 90%.

### 2. Testing Hypothesis by 3rd Decile <= 100
H0: decile.3 <= 100
H1: decile.3 > 100

#### a.) By Hand

```{r}
p <- 0.1
x <- 100
n <- length(weight)
t1 <- sum(weight <= x)

phi <- pbinom(t1, n, p)
phi
```
We actually get a very large p value asserting that the data supports our null hypothesis. We cannot reject it, using a 5% significance.

#### b.) Software Library
```{r}
binom.test(t1, n, p, "less")
```
We get a very close answer using the binom.test library function to our hand calculation using pbinom.  
Cannot reject the null hypothesis


## 4. Problem 4: Headroom
H0: 95th percentile = 70.3 inches
H1: 95th percentile ~= 70.3 inches
```{r}
heights <- c(72.6, 70.0, 71.3, 70.5, 70.8, 76.0, 70.1, 72.5, 71.1, 70.6, 71.9, 72.8)
```

### a.) By Hand
```{r}
t <- 70.3
p <- 0.95
n <- 100

t1 <- 100 - sum(heights > t)
t2 <- 100 - sum(heights >= t)

phi <- 2*min(pbinom(t1, n, p), 1 - pbinom(t2-1, n, p))
phi
```
Based on our approximation, we cannot reject the null hypothesis given a 5% significance level.

### b.) Software Library
```{r}
binom.test(t1, n, p, "two.sided")
binom.test(t2, n, p)
```
When we use the software library though, we fall within the 0.05 level and can reject the null hypothesis. 


## Problem 5: Tolerance Limits
Get the in class functions
```{r}
# translates from P( x^(r) <= at least q proportion of pop <= x^(r+m-1) ) <= alpha
bintol <- function(n, r, m, q)
  {
    i <- 0:(r+m-1)
    sum(choose(n, i) * (1-q)^i * q^(n-i))
  }

# Solves summation exactly
solve.bintol.n <- function(level, r, m, q)
  {
    ## convert from level to alpha
    alpha <- 1-level

    n <- 1
    while(1) {
        p <- bintol(n, r, m, q)
        if(p < alpha) break ## we crossed the threshold
        else n <- n+1
    }

    return(n)
  }

# Solves approximation by algebra
approx.bintol.n <- function(level, r, m, q)
  {
    x1ma <- qchisq(level, 2*(r+m))
    (1/4)*x1ma*(1+q)/(1-q) + (1/2)*(r+m-1)
}

solve.bintol.q <- function(level, n, r, m)
  {
    alpha <- 1-level
    f <- function(q, alpha, n, r, m) {
        alpha - bintol(n, r, m, q)
    }
    out <- uniroot(f, c(0,1), alpha=alpha, n=n, r=r, m=m)
    return(out$root)
}

approx.bintol.q <- function(level, n, r, m)
  {
      x1ma <- qchisq(level, 2*(r+m))
      temp <- 4*n - 2*(r+m-1)
      (temp - x1ma)/(temp + x1ma)
  }
```

### a.) Find sample size to be 90% confident that 95% of population will be within the range

#### i. Exactly
```{r}
r <- m <- 1
p <- 0.9
q <- 0.95
ans <- solve.bintol.n(0.9, 1, 1, 0.95)
ans
```

#### ii. Approximately
```{r}
approx.bintol.n(0.9, 1, 1, 0.95)
```
Which would be 76 if we rounded, or round up to 77 if we had the insight to be 'at least' 90% sure.

### b.) A fitness gym has measured the percentage of fat in 86 of its members.
```{r}
n <- 86
```

#### i. percentage between smallest and largest with 95% certainty
```{r}
r <- m <- 1
n <- 86
p <- 0.95

solve.bintol.q(p, n, r, m)
```

#### ii. what percentage between X(2) and X(85) with 95% certainty?
```{r}
r <- m <- 2
p <- 0.96
n <- 86
approx.bintol.q(p, n, r, m)
```

### c.) at least 90% of rods in a load are sixth longest/shortest with 99% confidence. Sample Size?
```{r}
r <- m <- 6
p <- 0.99
q <- 0.9
solve.bintol.n(p, r, n, q)
```

