---
title: "Homework 6"
author: "Michael Ahn"
date: "November 12, 2017"
output:
html_document:
  -smart
---
  
  ```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Problem 1: Temperature Revisited
```{r}
desmoines <- c(83, 91, 94, 89, 89, 96, 91, 92, 90)
spokane <- c(78, 82, 81, 77, 79, 81, 80, 81)
```

### a.) By Hand
H0: desmoines and spokane are identically distributed except diff means and Var(desmoines) >= Var(spokane)  
H1: var(desmoines) < var(spokane)  
```{r}
x <- desmoines
y <- spokane
xbar <- mean(desmoines)
ybar <- mean(spokane)
u <- abs(x - xbar)
v <- abs(y - ybar)

r <- rank(c(u, v))
r
ru <- r[seq_along(u)]
rv <- r[-seq_along(u)]
t <- sum(ru^2)
t

nx <- length(x)
ny <- length(y)
N <- nx + ny
r2bar <- (sum(ru^2) + sum(rv^2))/N
r4sum <- sum(r^4)
t1 <- (t - nx*r2bar)/sqrt(nx*ny/(N*(N-1))*r4sum - nx*ny/(N-1)*r2bar^2)
t1 

pnorm(t1, lower.tail=FALSE)
```
ranking, r: 17.0  2.5 15.0 10.5 10.5 16.0  2.5  9.0  4.0 12.0 13.0  7.0 14.0  5.0  7.0    1.0 7.0  
The test statistic is 0.8144833  
With a p-value of 0.2076841, we fail to reject the null hypothesis at a 5% significance level and conclude that the level of variability in the distributions of temperatures are equivalent  


### b.) Software Library
```{r}
#source("/Users/mahn/Desktop/School/STAT 3504/Functions/sqranks.R")
sqranks.test.approx <- function(x, y, alternative=c("two.sided", "greater", "less"))
{
  ## check for partial matches of the alternative argument
  alternative <- match.arg(alternative)
  
  ## calculate the absolute discrepancies from the (estimated) mean
  xbar <- mean(x)
  ybar <- mean(y)
  u <- abs(x - xbar)
  v <- abs(y - ybar)
  
  ## calculate the observed ranks, and t
  r <- rank(c(u, v))
  ru <- r[seq_along(u)]
  rv <- r[-seq_along(u)]
  t <- sum(ru^2)
  
  ## calculate the standardized version, t1
  nx <- length(x)
  ny <- length(y)
  N <- nx + ny
  r2bar <- (sum(ru^2) + sum(rv^2))/N
  r4sum <- sum(r^4)
  t1 <- (t - nx*r2bar)/sqrt(nx*ny/(N*(N-1))*r4sum - nx*ny/(N-1)*r2bar^2)
  
  ## calculate the p-value
  if(alternative == "two.sided") 
    phi <- 2*pnorm(-abs(t1))
  else if(alternative == "less")
    phi <- pnorm(t1)
  else if(alternative == "greater")
    phi <- pnorm(t1, lower.tail=FALSE)
  
  ## return the stats and p-value
  return(list(t=t, t1=t1, p.value=phi))
}
sqranks.test.approx(desmoines, spokane, "greater")
```
Same results. With a p-value of 0.2076841, we fail to reject the null hypothesis at a 5% significance level and conclude that the level of variability in the distributions of temperatures are equivalent.  



## Problem 2: Room Temperature Revisited
```{r}
men <- c(74, 72, 77, 76, 76, 73, 75, 73, 74, 75)
women <- c(75, 77, 78, 79, 77, 73, 78, 79, 78, 80)
```

### a.) By Hand
H0: men and women are identically distributed, possibly different means  
H1: the variance are not equal  
```{r}
x <- men
y <- women
xbar <- mean(x)
ybar <- mean(y)
u <- abs(x - xbar)
v <- abs(y - ybar)
r <- rank(c(u, v))
r #rank
ru <- r[seq_along(u)]
rv <- r[-seq_along(u)]
t <- sum(ru^2)
t #observed t
nx <- length(x)
ny <- length(y)
N <- nx + ny
r2bar <- (sum(ru^2) + sum(rv^2))/N
r4sum <- sum(r^4)
t1 <- (t - nx*r2bar)/sqrt(nx*ny/(N*(N-1))*r4sum - nx*ny/(N-1)*r2bar^2) 
t1 #standardized t
2*pnorm(-abs(t1))
```
ranking, r: 4.5 17.5 17.5 11.5 11.5 11.5  4.5 11.5  4.5  4.5 16.0  1.5  8.0 14.5  1.5 20.0   8.0 14.5  8.0 19.0  
test statistic: -0.7229738  
With a p-value of 0.469696, we fail to reject the null hypothesis and conclude that the variance among men and women are equal  

### b.) Software Library
```{r}
sqranks.test.approx(men, women, "two.sided")
```
Same results. With a p-value of 0.469696, we fail to reject the null hypothesis and conclude that the variance among men and women are equal  



## Problem 3: Rubber
```{r}
c20 <- c(2.07, 2.14, 2.22, 2.03, 2.21, 2.03, 2.05, 2.18, 2.09, 2.14, 2.11, 2.02)
c45 <- c(2.52, 2.15, 2.49, 2.03, 2.37, 2.05, 1.99, 2.42, 2.08, 2.42, 2.29, 2.01)
```

### a.) Numerical Summaries
```{r}
summary(c20)
```

```{r}
summary(c45)
```
We can see the quartiles, min, max, and means of the two samples  

### b.) Helpful Plot
```{r}
boxplot(c20, c45, names = c("c20", "c45"), xlab = "curing temp", ylab = "tensile strength")
```
We can see that the variability in c45 seems to be relatively much higher than for c20. Furthermore, the median for c45 seems to be close to the 3rd quartile for the c20 sample. The min and lower quartile are closer to one another, but the upper quartile for c45 is much higher; this implies that the mean for c45 will be higher than for c20, too.  

### c.) Two Sided Test comparing the mean
H0: the means for both temps are equal  
H1: the means are different  
```{r}
nx <- length(c20)
ny <- length(c45)
r <- rank(c(c20, c45))
r
rx <- r[seq_along(c20)]
t <- sum(rx) 
t
ties <- length(r) - length(unique(r))
ties
N <- nx + ny
r2sum <- sum(r^2)
r2sum
t1 <- t - nx*(N+1)/2
t1 <- t1/sqrt(nx*ny/((N*(N-1)))*r2sum - nx*ny*(N+1)^2/(4*(N-1)))
t1
2*min(pnorm(t1), pnorm(t1, lower.tail=FALSE))
```
With t = 130.5, and t1 = -1.12755  
With a p-value of 0.2595099, we fail to reject the null hypothesis at a 5% significance level and conclude that the means for both temperatures are the same  

#### d.) Two Sided Test comparing the variance
H0: the temps are identically distributed, potentially different means  
H1: the variances are different  
```{r}
sqranks.test.approx(c20, c45, "two.sided")
```
With an observed t = 857, and standardized t = -3.575  
With a p-value of 0.00035 we reject the null hypothesis at a 5% significance level and conclude that the variances between the two temperatures are different in their tensile strengths  



## Problem 4: Light Bulb Revisitted
```{r}
bulbs <- list(
    A=c(73, 64, 67, 62, 70), 
    B=c(84, 80, 81, 77),
    C=c(82, 79, 71, 75))
```
Can we conclude that the variability is all equal?  
H0: the three samples are identically distributed, possibly different means  
H1: there is at least one sample if a variance greater than another  
```{r}
sqranks.2plus.test <- function(xlist)
 {
    ## extract dimensions
    k <- length(xlist)
    n <- sapply(xlist, length)
    N <- sum(n)

    ## calculate absolute deviations from (estimated) mean and then rank them
    U <- lapply(xlist, function(x) { abs(x - mean(x)) })
    r <- rank(unlist(U))

    ## calculate s 
    g <- factor(rep.int(seq_len(k), n))
    s <- tapply(r, g, function(x) { sum(x^2) })

    ## calculate t2 via auxiliary quantities
    sbar <- sum(s)/N
    d2 <- (sum(r^4) - N*sbar^2)/(N-1)
    t2 <- (sum(s^2/n) - N*sbar^2)/d2

    ## finish with p-value
    phi <- pchisq(t2, k-1, lower.tail=FALSE)

    return(list(tstat=t2, p.value=phi))
}

sqranks.2plus.test(bulbs)
```
With a t-statitics of 1.055849 and p-value of 0.589828, we fail to reject the null hypothesis at a 5% significance level and conclude that the variance is equivalent for all the three bulb types.  



## Problem 5: Investments
```{r}
invest <- list(bonds=c(146, 180, 192, 185, 153),
    bluechip=c(176, 110, 212, 108, 196),
    speculative=c(-540, 1052, 642, -281, 67))

sqranks.2plus.test(invest)
```
H0: variability is identical  
H1: at least one of the variances among the 3 is not equal  
With a t-stat of 11.70985 and a p-value of 0.002865751, we reject the null hypothesis at the 5% significance level and conclude that the variance is significant for at least one of the investments  



## Problem 6: Couples' Bowling
```{r}
bowling <- data.frame(husband=c(147, 158, 131, 142, 183, 151, 196, 129, 155, 158),
    wife=c(122, 128, 125, 123, 115, 120, 108, 143, 124, 123))
```

### a.) Compute rho
```{r}
rho <- cor(rank(bowling$husband), rank(bowling$wife))
rho
```
rho = -0.6128049  

### b.) Two Tailed test for independence
H0: husband and wife scores are mutually independent  
H1: there tends to be a correlation relationship between small/large husband or small/large wife scores  
```{r}
cor.test(bowling$husband, bowling$wife, alternative="two.sided", method="spearman")
```
With a p-value of 0.05961 we fail to reject the null hypothesis at a 5% significance level and conclude that the scores are mutually independent (though it is very close)  

